---
title: "Advenced R: Envirionments and classes"
output: ioslides_presentation
---

```{r, include=FALSE}
library(ggplot2)
```

## Classes en R

* __S3__ (very useful)
* S4 (over engineered)
* RC (similar to Ruby's or Python's class, but R6 has some nicer
  features)
* __R6__ (very useful)
* proto (allows prototype-based programming)
* ggproto (similar to proto, used in ggplot)

## S3

```{r}
x <- data.frame(t=1:10, x=rnorm(10))
class(x)
```

```{r}
class(x) <- c("series", class(x))
class(x)
```

## S3

```{r}
plot(x)
```

## Method dispach

```{r}
plot.series <- function(series) {
  ggplot(series, aes(x=t, y=x)) + geom_line()
}
plot(x)
```

## New method

```{r}
average <- function(x) UseMethod("average")
average.numeric <- function(x) mean(x)
average(0:10)
```

```{r}
average.series <- function(x) average.numeric(x$x)
average(x)
```

## 

```{r}
add_to_series <- function(x, y) UseMethod("add_to_series")
add_to_series.series <- function(series_added, series) {
  df <- data.frame(t=series$t, x1=series$x, x2=series_added$x)
  class(df) <- c("few_series", class(df))
  df
}
y <- data.frame(t=1:10, x=runif(10))
class(y) <- c("series", class(y))

xy <- add_to_series(y, x)
head(xy, 2)
```

##

```{r}
add_to_few_series <- function(x, y) UseMethod("add_to_few_series")
add_to_few_series.series <-
  function(series_added, few_series) {
    n <- dim(few_series)[2]
    few_series[paste0("x", n)] <- series_added$x
    few_series
  }
head(add_to_few_series(x, xy), 2)
```
##

```{r}
add_to_few_series.few_series <-
  function(few_series_added, few_series) {
    m <- dim(few_series_added)[2]
    for (i in 1:(m-1)) {
      few_series <- add_to_few_series.series(
        data.frame(t=few_series_added$t,
                   x=few_series_added[paste0("x", i)]),
        few_series)
    }
    few_series
  }
head(add_to_few_series(xy, xy), 2)
```

##

```{r}
add_to_series.few_series <-
  function(few_series_added, series) {
    add_to_few_series.few_series(
      few_series_added,
      data.frame(t=series$t,
                 x1=series$x))
  }
head(add_to_series(xy, x), 2)
```
## Double dispach

We have defined 4 functions

* add_to_series.series
* add_to_series.few_series
* add_to_few_series.series
* add_to_few_series.few_series

Now we define `+`:
```{r}
add <- function(x, y) UseMethod("add")
add.series <- function(x, y) add_to_series(y, x)
add.few_series <- function(x, y) add_to_few_series(y, x)
```

## 

```{r}
head(add(x, y), 2)
```
```{r}
head(add(xy, x), 2)
```
```{r}
head(add(xy, xy), 2)
```

##

```{r}
class(x) <- c("seriess", class(x))
class(xy) <- c("seriess", class(xy))
`+.seriess` <- function(x, y) add(x, y)
head(x + x, 2)
```
```{r}
head(xy + x, 2)
```

## Environments

```{r}
dict <- new.env()
dict$a <- 1:5
dict$a
```

```{r}
dict[['d']] <- 20
dict$d
```
```{r}
assign('e', 40, envir=dict) 
dict$e
```



## Hidden variables and "ls"

```{r}
dict$.a <- 100
dict$.a
```

```{r}
ls(dict)
```


```{r}
ls(dict, all.names=TRUE)
```


## Extract values

```{r}
dict$a
```


```{r}
dict[['a']]
```

```{r}
get("a", dict)
```

## Parents

```{r}
dict2 <- new.env(parent=dict)
dict2$a
```

```{r}
get('a', dict2)
```

```{r}
parent.env(dict2)
dict
```
```{r}
dict2$f <- function() {
  print(environment())
  print(parent.env(environment()))
  
}

dict2$f()
```

## Object generator

```{r}
objectGenerator <- function(fields=list()) {
  self <- new.env()
  print(fields)
  for (fieldName in names(fields)) {
    self[[fieldName]] <- fields[[fieldName]]
  }
  self
}
```

```{r}
objectGenerator(list(time = 1:6, x = runif(6) ))
```

## R6

```{r}
library(R6)

Series6 <- R6Class(
  "Series6",
  public = list(
    time = NULL,
    x = NULL,
    initialize = function(time = NULL, x = NULL) {
      self$time <- time
      self$x <- x
    },
    plot = function() {
      df = data.frame(time=self$time, x=self$x)
      ggplot(df, aes(time, x)) + geom_line()
    }
  )
)
```

##

```{r}
x6 <- Series6$new(1:5, runif(5))
x6$plot()
```

##

```{r}
plot.Series6 <- function(nx) {nx$plot()}
plot(x6)
```




## Reactive Programming (fake code)

```
x <- 2
y <- 2 * x
y
```
4
```
x <- 4
y
```
8

## Reactive Programming (real code)

```{r}
input <- new.env()
output <- new.env()
parent.env(output) <- input
class(output) <- "reactive"

`$.reactive` <- function(e, v) {
  v_name <- as.character(substitute(v))
  get(v_name, envir=e)()
}

reactive <- function(block) {
  block_name <- as.character(substitute(block))[2]
  function() eval(parse(text=block_name))
}
```

##
```{r}
input$x <- 10
input$y <- 20
output$z <- reactive({input$x + input$y})
output$z
input$x <- 17
output$z
```

##
```{r}
output$w <- reactive({output$z + input$y})
output$w
input$x <- 0
output$w
```